Quantitative Evaluation of Touch-Screen Performance with Transition Matrix Measures

### 2015-08-27 flux-entropy-paper/flux-entropy-paper.tex
new calculations for other matrix measures

In the following sections, we will use transition matrices, flux, and
entropy to analyse a corpus of data collected as part of our research
into collaborative interactions with musical touch-screen tablet
interfaces. In a period of more than two years (April 2013 - June 2015), we have
collected data from $139$ sessions where various participants used
musical touch-screen interfaces on Apple iPad devices. Each session
took place in a number of contexts including rehearsals, performances, lab study
sessions, classes, and casual demonstrations. Participants performed
free improvisations as well as written compositions and performed on
iPads alone or using setups of iPads and acoustic percussion
instruments. The median length of these sessions was 7m32s (minimum was
1m0s and maximum was 26m) and the
median number of participants was four (minimum 1, maximum 9).

>>>>

In the following sections, we will use transition matrices, flux, and
entropy to analyse a corpus of data collected as part of our research
into collaborative interactions with musical touch-screen tablet
interfaces. In a period of more than two years (April 2013 - June 2015), we have
collected data from $150$ sessions where various participants used
musical touch-screen interfaces on Apple iPad devices. Each session
took place in a number of contexts including rehearsals, performances, lab study
sessions, classes, and casual demonstrations. Participants performed
free improvisations as well as written compositions and performed on
iPads alone or using setups of iPads and acoustic percussion
instruments. The median length of these sessions was 7m28s (minimum was
1m0s and maximum was 26m) and the
median number of participants was four (minimum 1, maximum 7).

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figures/sessions-count}
  \caption{
    \label{fig:count-data}}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figures/flux-boxplot}
  \caption{
    \label{fig:flux-boxplot}}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figures/entropy-boxplot}
  \caption{
    \label{fig:entropy-boxplot}}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figures/flux-entropy-distribution}
  \caption{
    \label{fig:flux-entropy-distribution}}
\end{figure}

A three-way ANOVA procedure was performed on the data set to find
whether performance context, type, and instrumentation could predict
the entropy and flux values.

Using performative sessions only (i.e. not the demonstration
sessions). We found a significant main effect on flux due to
performance type ($F(1,86) = 5.87, p < 0.05$), and interaction
effect due to performance type and instrumentation ($F(1,86) = 12.37, p
< 0.001$). Similar effects on entropy were found due to performance
type ($F(1,86) = 13.62, p < 0.001$) and an interaction effect due to
performance type and instrumentation ($F(1,86) = 4.98, p < 0.05$). The
effect on entropy due to performance context (rehearsal,
performance, or study) was not found to be significant at $\alpha =
0.05$, ($F(2,86) = 2.11, p = 0.128$). When including demonstration
sessions in the analysis, however, performance context was found to
have a significant main effect on entropy ($F(4,138) = 5.07, p <
0.001$).
